# Chi-Square Tests

*July 26, 2023*

1.  The chi-square test is a statistical test used to determine if there is a significant difference between the **observed values** and the **expected values** in a categorical data set.
2.  It measures the **deviation** between the expected and observed frequencies in one or more categories and assesses whether this deviation is statistically significant.
3.  The result of a chi-square test is a test statistic, and its p-value is compared against a threshold (e.g. $\alpha = 0.05$) to determine the statistical significance of the observed differences.
4.  The test is commonly used in **hypothesis testing**, contingency table analysis, and **goodness-of-fit** testing.

## Types of Tests

1.  There are several types of chi-square tests. The most popular tests are as follows.

| Test                        | Use                                                                        |
|-----------------------|-------------------------------------------------|
| **A. Goodness-of-fit test** | determine if a sample of data **fits a specified distribution**            |
| **B. Independence test**    | determine if there is a **relationship between two categorical variables** |

2.  Additional chi-square tests include:

-   Homogeneity test: used to determine if different populations have the same distribution of a categorical variable.
-   Contingency table test: used to analyze the relationship between two or more categorical variables in a multi-dimensional table.
-   McNemar's test: used to determine if the difference between paired nominal data is significant.
-   Likelihood-ratio test: used to compare nested models, where the more complex model is tested against a simpler model.
-   Mantel-Haenszel test: used to determine if there is a relationship between two categorical variables while controlling for the effect of a third variable.

## Chi-Square Goodness of Fit test

### A Business Application of the Chi-Square Goodness of Fit test

1.  Suppose a **retail** company wants to know **if the gender distribution of their customer base is representative of the general population**. They collect data on the gender of a sample of their customers and compare it to the expected distribution (e.g. 50% male and 50% female).
2.  The Chi-Square Goodness of Fit test is then used to evaluate the **null hypothesis** that the observed distribution of gender among the company's customers is the same as the expected distribution.
3.  The test could help us potentially conclude that there is a significant difference between the observed and expected distributions, and the company's customer base may not be representative of the general population.
4.  For example, **the results might show that a significantly higher proportion of females than expected are customers of the company**. This information could be used by the retail company to tailor their marketing strategies and product offerings to better attract male customers.

### Chi-Square Goodness of Fit test -- Technicalities

1.  The chi-square goodness-of-fit test is a statistical test that is used to determine if a sample of categorical data fits a specified distribution.
2.  It compares the **observed frequencies** of different categories in the sample data to the **expected frequencies** based on the specified distribution.
3.  The test statistic is calculated as the sum of the squared differences between the observed and expected frequencies, divided by the expected frequencies.
4.  **If the test statistic is larger than the critical value from a chi-square distribution table**, it can be concluded that there is a significant difference between the observed and expected distributions
5.  The resulting p-value is then compared against a threshold (e.g. alpha=0.05) to determine the statistical significance of the observed differences.
6.  It can be used to test hypotheses about the distribution of a categorical variable, or testing if a population is homogeneous with respect to a particular attribute.
7.  The test assumes that the sample size is large enough and that the expected frequencies are greater than or equal to 5 for all categories.

### Running the Chi-Square Goodness of Fit Test in R

1.  The `mtcars` dataset consists of 32 cars, each having 3, 4 or 5 gears.

```{r}
# Load the data
data(mtcars)

# Create a table of the number of cars in each gear category
t0 <- table(mtcars$gear)
t0
```

2.  Graphical display of Contingency table:

```{r echo=TRUE, eval=TRUE}
library("gplots")
# 1. convert the data as a table
dt <- as.table(as.matrix(t0))
# 2. Graph
balloonplot(t(dt), main ="Contingency Table",
            xlab ="gear",
            label = FALSE, show.margins = FALSE)
```

3.  Proportions

```{r}
t1 = prop.table(t0)
t1
```

4.  Running the test

```{r }
# Specify the distribution we want to test against
expected_probs <- c(0.5, 0.3, 0.2)

# Perform the chi-square goodness-of-fit test
chisq.test(t1, p=expected_probs)

```

5.  This code creates a table of the number of cars in each gear category, specifies the distribution we want to test against (in this case, a distribution where 50% of the cars have 3 gears, 30% have 4 gears, and 20% have 5 gears), calculates the expected frequencies based on the expected probabilities, and then performs the chi-square goodness-of-fit test.
6.  The resulting p-value will indicate the significance of the observed differences between the observed and expected frequencies.

## Chi-Square Test of Independence

### A Business Application of the Chi-Square Test of Independence

1.  Suppose a grocery store wants to evaluate the **association** between the **type of product** a customer buys and their **age group**.
2.  Suppose the grocery store wants to know if there is a significant association between the **type of product a customer buys** (**e.g. fruits, vegetables, or dairy products**) and their **age group** (e.g. **18-30, 31-45, 46-60, 61 and older**). They collect data on the age group and product type for a sample of customers and create a contingency table.
3.  The Chi-square test is then used to evaluate the **null hypothesis** that the **product type and age group are independent**.
4.  Depending on the test result, we could potentially conclude that there is a significant association between the two variables.
5.  For example, the results might show that **a significantly higher proportion of customers in the 46-60 age group buy fruits compared to the other age groups**.
6.  This information could be used by the grocery store to adjust their marketing strategies and product offerings to better cater to their target customers.

### Chi-Square Test of Independence -- Technicalities

1.  The Chi-square test of independence is a statistical test used to determine if there is a significant association between two categorical variables.
2.  The test evaluates the **null hypothesis** that the **two variables are independent** and **calculates a test statistic (chi-square)** based on the difference between the observed and expected frequencies of the two variables.
3.  If the calculated test statistic is larger than the **critical value from a chi-square distribution table**, then the null hypothesis is rejected and it can be concluded that there is a significant association between the two variables.

### Running the Chi-Square Test of Independence Test in R

1.  Convert the categorical variables into factor variables in the dataset mtcars

```{r echo=TRUE, eval=TRUE}
data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$am  <- as.factor(mtcars$am)
mtcars$gear  <- as.factor(mtcars$gear)
```

2.  Creating a Contingency Table:

```{r echo=TRUE, eval=TRUE}
ctab <- table(mtcars$am, mtcars$cyl)
ctab
```

3.  Graphical display of Contingency table:

```{r echo=TRUE, eval=TRUE}
library("gplots")
# 1. convert the data as a table
dt <- as.table(as.matrix(ctab))
# 2. Graph
balloonplot(t(dt), main ="Contingency Table",
            xlab ="cyl", ylab="am",
            label = FALSE, show.margins = FALSE)
```

4.  Compute Chi-Square test: The Chi-square statistic can be easily computed using the function `chisq.test()` as follow:

```{r echo=TRUE, eval=TRUE}
chisq <- chisq.test(ctab)
chisq
```

5.  In our example, the row and the column variables are statistically significantly associated (p-value = 0.01265).

6.  The observed and the expected counts can be extracted from the result of the test as follows:

```{r echo=TRUE, eval=TRUE}
# Observed counts
chisq$observed

# Expected counts
round(chisq$expected, 2)
```

7.  Pearson Residuals: **Positive residuals** are positive values in cells specify an attraction (positive association) between the corresponding row and column variables. **Negative residuals** implies a repulsion (negative association) between the corresponding row and column variables.

```{r echo=TRUE, eval=TRUE}
round(chisq$residuals, 3)
```
